---
title: "Insect survey 2020, basic analysis"
author: "Jens Å"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---

**To-do**
* Need to compare vaatvekt minus bottles. Or better, tørrvekt.


* These species are not matched to the gyldig names in artsdatabasen yet.
* Should store dates and times separately as well in logger_data table. Easier to subset later.
* Update weight figures with bottle weight removed.
* Notera i rapport hur många treffer som er misstenkte.  


* Make plots of phenology
  - No. species over the season for each location. Split into separate figs for the two habitats?
  - Relate this to local temperature, temperature sums?
  
* Plots of number of samples over time

* Check differences btw 2 week and 4 week sampling
  - Species occurrences, 
  - compare 2 * 2 weeks vs. 1 * 4 week
  - Compare first and second of the 2 weeks, see if any of the have more overlap with 4 week. If 4 week becomes full, the first 2 week period should overlap more.
  - Do this for window traps as well.
* Check overlap/additions of window traps vs. malaise traps
  - Subset only Coleoptera
  - Group by location + time
* Within habitat, check relationship btw species richness and a few variables
  - Forest age
  - Diversity of fauna
  - Temperature
  - Altitude
* Check differences btw ethanol and propglyk window traps

```{r, include = F}
#Some common packages, loading rmarkdown doesn't like the messages from tidyverse, so we don't include this in the document'
require(tidyverse)
require(DBI)
require(RPostgreSQL)
require(dbplyr)
require(ggplot2)
require(xtable)
require(NinaR)
require(gridExtra)
require(hms)
require(ggeffects)
require(lme4)
```


```{r setup, include=FALSE}
#This is optional
#I choose the 'styler' package for tidying the code to preserve indentations
#I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
#Set tidy = True to get the knitr default
#I want all figures as png and pdf in high quality in a subfolder called figure 

knitr::opts_chunk$set(echo = TRUE, 
                      tidy = "styler",
                      dev = c("png", "pdf"),
                      dpi = 600,
                      fig.path = "figure/",
                      tidy.opts = list(width.cutoff = 60)
                      )

options(xtable.comment = F, 
        xtable.include.rownames = F, 
        nina.logo.y.pos = 0.15)
palette(ninaPalette())
```

Data import and cleaning
==============
At this point I get the raw data from Marie, which requires some cleaning. In the future, we'll put this all in a database, and separate out the cleaning and database import in a separate script.

```{r, include = F, eval = T}
source("~/.rpgpass")
#This connects to the gisdatabase with a DBI connection named `con`.
#Use for example dbGetQuery(con, "SELECT * FROM ....") to query the database
con <- dbConnect(PostgreSQL(),
          host = "ninradardata01.nina.no", 
                 dbname = "insect_monitoring",
                 user = username,
                 password = password
)
rm(username, password)
```


```{r import_csv}

#dat <- read.delim("../../Data/Genetikkdata/long_format_data_GCF-2020-735_736_19102020.txt", sep= " ") %>% 
#  as.tibble()

#dat <- read_delim("../../Data/Genetikkdata/long_format_data_GCF-2020-735_736_19102020.txt",

dat <- read_delim("../../Data/Genetikkdata/long_format_data_GCF-2020-735_736_744_26102020.txt", 
                  delim= ";",
                  col_types = cols(
  kingdom = col_character(),
  phylum = col_character(),
  class = col_character(),
  order = col_character(),
  family = col_character(),
  genus = col_character(),
  species = col_character(),
  sample = col_character(),
  value = col_double(),
  GenlabID = col_character(),
  GUID = col_character(),
  Tilsatt_antall_C.analis_SI_USA_08.06.20 = col_double(),
  Tilsatt_antall_C.chinensis_08.06.20 = col_double(),
  Tilsatt_antall_Sirisser = col_double(),
  Tilsatt_antall_Melormer = col_double(),
  Tilsatt_antall_C.maculatus = col_double(),
  Vekt_tom_flaske_.g. = col_double(),
  Vekt_flaske_med_EtOH_og_insekter_.g. = col_double(),
  Vekt_flaske_._insekter_vaat_.g. = col_double(),
  Vekt_flaske_._insekter_toerr_.g. = col_double(),
  Totalvolum_.mengde_ATL_._prot.K_mL. = col_character(),
  Vaatvekt_.g. = col_double(),
  Toerrvekt_.g. = col_double(),
  Kommentar_proeve = col_character(),
  Ekstraksjonsdato = col_date(format = "%d.%m.%Y"),
  Ekstraksjonskit = col_character(),
  Subsamplet_volum_.ul. = col_double(),
  Elueringsvolum = col_double(),
  OD = col_double(),
  X260_280 = col_double(),
  X260_230 = col_double(),
  Labperson = col_character(),
  Ekstraksjonskommentar = col_character(),
  locality = col_character(),
  trap = col_character(),
  date_placed = col_date(format = "%d.%m.%Y"),
  time_placed = col_time(format = ""),
  empty_week = col_double(),
  date_emptied = col_date(format = "%d.%m.%Y"),
  time_emptied = col_time(format = ""),
  liquid = col_character(),
  person_emptied = col_character(),
  sample_id = col_character(),
  mottat_lab = col_character(),
  comment = col_character()
),
na = c("NA",
       "N/A",
       "Maa veies"
)
)


```
```{r, eval = F, echo = F}
spec(dat)
```

Fix some coding errors.
```{r fix_coding}
# dat %>% 
#   select(genus, species, new_species) %>% 
# slice(297, 1902)
#   
# 
# dat %>% 
#   select(genus, species, new_species) %>% 
# slice(3219, 3585, 4017, 11424, 13275, 15465, 21056, 24168, 25077, 31567, 32805, 48502, 49967)
#   
#Ekstraksjonskommentar = gsub(pattern = "[\\\"]", "", Ekstraksjonskommentar),

dat <- dat %>% 
  mutate(species_latin = species)

dat <- dat %>% 
  mutate(species = gsub(" sp", "_sp", species))

dat <- dat %>% 
  mutate(species = gsub("([0-9])(_)(sp)", "\\1/\\3", species))

dat <- dat %>%
  mutate(species = gsub("(.*)(_)(.*)", "\\3", species))

dat <- dat %>% 
  mutate(locality = str_to_sentence(locality))

```

Set factor levels

```{r}
dat <- dat %>%
  mutate(locality = factor(locality, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10), "Nerlandsoeya")),
         trap_type = ifelse(grepl("MF", trap), "Malaise", "Vindu"),
         habitat_type = ifelse(grepl("Skog", locality), "Skog", "SN_Gressmark"),
         liquid = ifelse(grepl("etoh85", liquid), "Ethanol", "PropGl_Ethanol"),
         weeks_sampled = ifelse(grepl("1", trap) | grepl("3", trap), 2, 4)
         )
```

Write a species list for the entomologists
--------

```{r}
dat %>% 
  select(kingdom,
         phylum,
         class,
         order,
         family,
         genus,
         species) %>% 
  distinct() %>% 
  arrange(kingdom,
         phylum,
         class,
         order,
         family,
         genus,
         species) %>% 
  write_excel_csv(path = "out/species_list_2020-11-02.csv")
```



Filter out the Nerlandsøya samples
===========
```{r}
dat <- dat %>% 
  filter(locality != "Nerlandsoeya") %>% 
  droplevels()
```

Add missing bottle weights. This is based on a sample of bottle dry weight measurements made after the regular weighings. 

```{r}
bottle_weights <- tbl(con,
                   in_schema("catch_data", "bottle_weights")) %>% 
  select(-id)


bottle_weights %>% 
  filter(bottle_type == "MF" & year == 2000) %>% 
  select(dry_weight) %>% 
  ggplot(.) +
  geom_density(aes(x = dry_weight))

bottle_weights_MF <- bottle_weights %>% 
  filter(bottle_type == "MF" & year == 2000) %>% 
  select(dry_weight) %>% 
  summarize(mean_weight = mean(dry_weight)) %>% 
  collect()


```

```{r}
sum(is.na(dat$Vekt_tom_flaske_.g.))
dat <- dat %>% 
  mutate(Vekt_tom_flaske_.g. = ifelse(trap_type == "Malaise" & is.na(Vekt_tom_flaske_.g.), bottle_weights_MF$mean_weight, Vekt_tom_flaske_.g.))
   
sum(is.na(dat$Vekt_tom_flaske_.g.))
```

```{r}
dat <- dat %>% 
  mutate(Vaatvekt_.g. = Vekt_flaske_._insekter_vaat_.g. - Vekt_tom_flaske_.g.,
         Toerrvekt_.g. = Vekt_flaske_._insekter_toerr_.g. - Vekt_tom_flaske_.g.)
```



Check the emptying dates.
----------------
The dates seems to be OK this time. Should do a proper check on the endpoints though.
```{r}
dat <- dat %>% 
  mutate(days_collected = as.numeric(dat$date_emptied - dat$date_placed))

# dat %>% 
#   filter(days_collected == 0) %>% 
#   select(sample_id,
#          GUID,
#          date_placed,
#          date_emptied) %>% 
#   distinct() %>% 
#   print(n = Inf)
# 
# hist(as.numeric(dat$date_emptied - dat$date_placed))
```

<!-- These traps have erroneous placement dates. -->
<!-- ```{r} -->
<!-- dat %>%  -->
<!--   filter(sample_id == "skog_1_MF1_week_") %>%  -->
<!--   slice(1:2) %>%  -->
<!--   print(width = Inf) -->

<!-- dat %>%  -->
<!--   filter(sample_id == "skog_1_MF1_week_26") -->
<!-- ``` -->

```{r}

dat %>% 
  group_by(days_collected) %>% 
  summarise(n())

# dat %>% 
#   filter(days_collected == 8) %>% 
#   select(sample_id,
#          GUID,
#          date_placed,
#          date_emptied) %>% 
#   distinct() %>% 
#   print(n = Inf)
```

Set weeks sampled manually for the few locations in Oslo that got an extra short sampling in the beginning.
```{r}
ggplot(dat, aes(x = weeks_sampled, y = days_collected, group = weeks_sampled)) +
  geom_boxplot()

```


```{r}
twoWeeks <- dat %>% 
  filter(weeks_sampled == 4 &
           days_collected < 20) %>% 
  select(sample_id) %>% 
  distinct() %>% 
  as.vector()


dat <- dat %>% 
  mutate(weeks_sampled = ifelse(sample_id %in% twoWeeks[[1]], 2, weeks_sampled))


```

```{r}
ggplot(dat, aes(x = weeks_sampled, y = days_collected, group = weeks_sampled)) +
  geom_boxplot()

```

 
Check missing weight data
===========
```{r}
dat %>% 
  group_by(empty_week) %>% 
  select(Vekt_tom_flaske_.g.,
         Vekt_flaske_med_EtOH_og_insekter_.g.,
         Vekt_flaske_._insekter_vaat_.g.,
         Vekt_flaske_._insekter_toerr_.g.,
         Vaatvekt_.g.,
         Toerrvekt_.g.
         ) %>%  
  distinct() %>% 
  summarise_all(~sum(is.na(.)))
```
 
 
 
 

Basic species richness graphs
==============

Quick look at diversity at the different sites.

```{r}
total_no_taxa <- dat %>% select(order, family,genus,species) %>% 
  distinct() %>% 
  count()
```


```{r no_spec_loc_malaise, fig.cap = "Cumulative number of taxa found in malaise traps."}
dat %>% 
  filter(grepl("MF", trap)) %>% 
  group_by(locality) %>% 
  select(genus,species) %>% 
  distinct() %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(.) +
 geom_bar(aes(y = n, x = locality, fill = locality),
           stat = "identity") +
 # ggtitle(paste0("Kumulativ artsrikedom i\nalle malaisefeller per ", max(dat$date_emptied))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") +
    scale_fill_nina() +
  ylab("Antall taksa") +
  xlab("Lokalitet")
```

```{r no_spec_loc_window, fig.cap = "Cumulative number of taxa found in window traps."}
dat %>% 
  filter(grepl("VF", trap)) %>% 
  group_by(locality) %>% 
  select(genus,species) %>% 
  distinct() %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(.) +
  geom_bar(aes(y = n, x = locality, fill = locality),
           stat = "identity") +
 #  ggtitle(paste0("Kumulativ artsrikedom i\nalle vindusfeller per  ", max(dat$date_emptied))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none")  +
    scale_fill_nina() +
  ylab("Antall taksa") +
  xlab("Lokalitet")
```



Basic biomass graphs
==============
NB, we still have some missing data on the weight of the bottles.

Quick quality check of the spread of the bottle weights. Could we impute the missing values?
```{r}
dat %>% 
  select(Vekt_tom_flaske_.g.,
         Vekt_flaske_med_EtOH_og_insekter_.g.,
         Vekt_flaske_._insekter_vaat_.g.,
         Vekt_flaske_._insekter_toerr_.g.,
         Vaatvekt_.g.,
         Toerrvekt_.g.)
```

```{r}
ggplot(dat)  +
  geom_histogram(aes(Vekt_tom_flaske_.g., fill = trap_type))
```


```{r}
hist(dat$Vekt_tom_flaske_.g.)
```

```{r}
table(dat$Vekt_tom_flaske_.g.)
```
These bottle weights are weirdly consistent. Did they really weigh all bottles? UPDATE WITH NEW BOTTLE WEIGHTS. !!DONE!!

```{r}
summary(dat$Toerrvekt_.g.)
```



Quick look at the biomass at the different sites.

```{r dry_weight_loc_malaise, fig.cap = "Average biomass (dry weight) in malaiase traps."}
dat %>% 
  filter(grepl("MF", trap)) %>% 
  group_by(locality) %>% 
  ggplot(.) +
 geom_boxplot(aes(y = Toerrvekt_.g., x = locality, fill = locality)) +
 # ggtitle(paste0("Middels avrunnen vekt (pluss flaske)\nav insekter i malaisefeller per ", max(dat$date_emptied))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") +
    scale_fill_nina() +
  ylab("Vekt (g.)") +
  xlab("Lokalitet")
```

```{r dry_weight_loc_window, fig.cap = "Average biomass (dry weight) in window traps."}
dat %>% 
  filter(grepl("VF", trap)) %>% 
  group_by(locality) %>% 
  ggplot(.) +
 geom_boxplot(aes(y = Toerrvekt_.g., x = locality, fill = locality)) +
  #ggtitle(paste0("Middels avrunnen vekt (pluss flaske) av insekter i\nvindusfeller per ", max(dat$date_emptied))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") +
    scale_fill_nina() +
  ylab("Vekt (g.)") +
  xlab("Lokalitet")
```



Check overlap between malaise and window traps
==============
Simple: total spec richn for all traps and times per location, plus the same for just malaise and window traps. group in figure by location

complex: tally how many spec is found in only one trap type. No unecessary

per species: which sp is only found in one trap type. or per species, % of findings in each trap type

Total number of coleoptera species per trap type
-----------

```{r}
col_spec_richn_tot <- dat %>% 
  filter(order == "Coleoptera",
         habitat_type == "Skog") %>% 
  group_by(locality) %>% 
  summarize(col_spec_richn = n_distinct(genus, species)) %>% 
  mutate(trap_type = "Alle")
  
```
```{r}
col_spec_richn_trap <- dat %>% 
  filter(order == "Coleoptera",
         habitat_type == "Skog") %>% 
  group_by(locality,
           trap_type) %>% 
  summarize(col_spec_richn = n_distinct(genus, species))
  
```

```{r}
col_spec_richn <- col_spec_richn_tot %>% 
  union_all(col_spec_richn_trap)
```

```{r col_spec_richn_trap, fig.cap = "Cumulative number of Coleoptera taxa identified in each trap type"}
ggplot(col_spec_richn, aes(y = col_spec_richn, 
                           x = locality, 
                           fill = trap_type)) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
 #   ggtitle(paste0("Kumulativt antall billearter (identifiserte taksa)\ni hver felletyp per ", max(dat$date_emptied))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_nina(name = "Felletype") +
  ylab("Antall taksa") +
  xlab("Lokalitet")
  
```

Percentage of catches in each trap type per species
----------
How to make it more readable? Split out all the singletons into a separate graph?

```{r}
col_perc_traptype <- dat %>% 
  filter(order == "Coleoptera",
         habitat_type == "Skog") %>% 
  group_by(trap_type,
           genus,
           species_latin) %>% 
  summarize(no_catches = n())
  
```

```{r col_perc_traptype, fig.width = 12}

p1 <- col_perc_traptype %>% 
  ungroup() %>% 
  arrange(species_latin,
          trap_type) %>% 
  filter(no_catches > 1) %>% 
    slice(1:50) %>% 
  group_by(trap_type) %>% 
ggplot(.) +
  geom_bar(aes(y = reorder(species_latin, desc(species_latin)), x = no_catches, fill = trap_type), 
           stat = "identity",
           position = position_dodge()) +
  scale_fill_nina(name = "Trap type") +
  ylab("") +
  xlab("Number of catches") +
   theme(legend.position = "none",
         text = element_text(size=5))


p2 <- col_perc_traptype %>% 
  ungroup() %>% 
  arrange(species_latin,
          trap_type) %>% 
  filter(no_catches > 1) %>% 
    slice(51:100) %>% 
  group_by(trap_type) %>% 
ggplot(.) +
  geom_bar(aes(y = reorder(species_latin, desc(species_latin)), x = no_catches, fill = trap_type), 
           stat = "identity",
           position = position_dodge()) +
  scale_fill_nina(name = "Trap type") +
  ylab("") +
  xlab("Number of catches") +
   theme(legend.position = "none",
         text = element_text(size=5))

p3 <- col_perc_traptype %>% 
  ungroup() %>% 
  arrange(species_latin,
          trap_type) %>% 
  filter(no_catches > 1) %>% 
    slice(101:150) %>% 
  group_by(trap_type) %>% 
ggplot(.) +
  geom_bar(aes(y = reorder(species_latin, desc(species_latin)), x = no_catches, fill = trap_type), 
           stat = "identity",
           position = position_dodge()) +
  scale_fill_nina(name = "Trap type") +
  ylab("") +
  xlab("Number of catches") +
   theme(legend.position = "none",
         text = element_text(size=5))


p4 <- col_perc_traptype %>% 
  ungroup() %>% 
  arrange(species_latin,
          trap_type) %>% 
  filter(no_catches > 1) %>% 
    slice(151:200) %>% 
  group_by(trap_type) %>% 
ggplot(.) +
  geom_bar(aes(y = reorder(species_latin, desc(species_latin)), x = no_catches, fill = trap_type), 
           stat = "identity",
           position = position_dodge()) +
  scale_fill_nina(name = "Trap type") +
  ylab("") +
  xlab("Number of catches") +
   theme(text = element_text(size=5))

grid.arrange(p1, p2, p3, p4, nrow = 1)

```


```{r no_col_singletons}
col_perc_traptype %>% 
  ungroup() %>% 
  arrange(species_latin,
          trap_type) %>% 
  filter(no_catches ==1) %>% 
  group_by(trap_type) %>% 
  summarize(no_singletons = n()) %>% 
 ggplot(.) +
  geom_bar(aes(y = no_singletons, x = trap_type, fill = trap_type), 
           stat = "identity") +
  scale_fill_nina(name = "Trap type") +
  ylab("Number of singleton Coleoptera species") +
  xlab("")
```



Figs to Niclas on Pollinators
============
Niclas Gyllenstrand asked how efficient the malaisetraps+metabarcoding are at finding pollinators. This could be a relevant graph for the report as well.



```{r}
poll_spec_richn <- dat %>% 
  filter(trap_type == "Malaise") %>% 
  filter(family == "Andrenidae" |
         family == "Apidae" |
         family == "Colletidae" |
         family == "Halictidae" |
         family == "Megachilidae" |
         family == "Melittidae" |
         family == "Syrphidae" |
          
         family == "Hesperiidae" |
         family == "Lycaenidae" |
         family == "Nymphalidae" |
         family == "Papilionidae" |
         family == "Pieridae" |
         family == "Riodinidae"
           ) %>% 
  group_by(family) %>% 
  distinct(family, genus, species) %>% 
  summarize(col_spec_richn = n(),
            no_unidentified_sp = sum(grepl("sp.", species)))
```


```{r poll_spec_richn}
ggplot(poll_spec_richn, aes(x = family, y = col_spec_richn, label = col_spec_richn)) +
  geom_bar(stat = "identity", 
           aes(fill = family)) + 
  geom_text(size = 3,
            position = position_nudge(y = 4)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_fill_nina(name = "Family") +
  xlab("") +
  ylab("Number of species")
```


```{r}
pollinator_species <- dat %>% 
  filter(trap_type == "Malaise") %>% 
  filter(family == "Andrenidae" |
         family == "Apidae" |
         family == "Colletidae" |
         family == "Halictidae" |
         family == "Megachilidae" |
         family == "Melittidae" |
         family == "Syrphidae" |
          
         family == "Hesperiidae" |
         family == "Lycaenidae" |
         family == "Nymphalidae" |
         family == "Papilionidae" |
         family == "Pieridae" |
         family == "Riodinidae"
           ) %>% 
  distinct(order, family, genus, species) %>% 
  arrange(family, genus, species) 
```


```{r}
dat %>% 
  filter(trap_type == "Malaise") %>% 
  filter(family == "Apidae") %>% 
  distinct(genus, species) %>% 
  select(genus, species)
```

```{r}
write_csv(pollinator_species, 
          path = "out/pollinator_species.csv")
```





Phenology graphs
===========
Some "timeseries" throughout the season. 

```{r}
seasonAvgNo <- dat %>% 
  filter(weeks_sampled == 2) %>% 
  group_by(habitat_type, date_emptied, locality) %>% 
  summarise(no_species = n_distinct(genus, species),
            sum_biomass = sum(Vekt_flaske_._insekter_vaat_.g., na.rm = T) / no_species)


# test <- seasonAvgNo <- dat %>% 
#   filter(weeks_sampled == 2) %>% 
#   group_by(habitat_type, date_emptied, locality) %>% 
#   select(habitat_type, date_emptied, locality, Vekt_flaske_._insekter_vaat_.g.) %>% 
#   distinct() %>% 
#   summarise(sum(Vekt_flaske_._insekter_vaat_.g., na.rm = T))

```

```{r no_spec_phenology_forest}
seasonAvgNo %>% 
  filter(habitat_type == "Skog") %>% 
ggplot() +
  geom_point(aes(x = date_emptied, y = no_species, color = locality)) +
  geom_line(aes(x = date_emptied, y = no_species, color = locality)) +
  ylab("Antall taksa") +
  xlab("Tømmedato") +
  scale_color_nina(name = "Lokalitet")
```

```{r no_spec_phenology_grassland}
seasonAvgNo %>% 
  filter(habitat_type == "SN_Gressmark") %>% 
ggplot() +
  geom_point(aes(x = date_emptied, y = no_species, color = locality)) +
  geom_line(aes(x = date_emptied, y = no_species, color = locality)) +
  ylab("Antall taksa") +
  xlab("Tømmedato") +
  scale_color_nina(name = "Lokalitet")
```

```{r biomass_phenology_forest}
seasonAvgNo %>% 
  filter(habitat_type == "Skog") %>% 
ggplot() +
  geom_point(aes(x = date_emptied, y = sum_biomass, color = locality)) +
  geom_line(aes(x = date_emptied, y = sum_biomass, color = locality)) +
  ylab("Total innsamlet biomasse (våtvekt g., pluss flaske)") +
  xlab("Tømmedato") +
  scale_color_nina(name = "Lokalitet")
```

```{r biomass_phenology_grassland}
seasonAvgNo %>% 
  filter(habitat_type == "SN_Gressmark") %>% 
ggplot() +
  geom_point(aes(x = date_emptied, y = sum_biomass, color = locality)) +
  geom_line(aes(x = date_emptied, y = sum_biomass, color = locality)) +
  ylab("Total innsamlet biomasse (våtvekt g. pluss flaske)") +
  xlab("Tømmedato") +
  scale_color_nina(name = "Lokalitet")
```


Plot some weather data
=======


```{r}
logger_tbl <- tbl(con,
                   in_schema("loggers", "logger_data")) %>% 
  select(-id)

logger_dep_tbl <- tbl(con,
                   in_schema("loggers", "logger_deployments")) %>% 
  select(-id)
```

```{r}
logger_data <- logger_tbl %>% 
  left_join(logger_dep_tbl,
            by = c("logger_id" = "logger_id",
                   "logger_type" = "logger_type")) %>% 
  mutate(day = as.Date(date))
```

Plot average values for each site during July

```{r avg_logger_july}
tt <- logger_data %>% 
  group_by(location, day, data_type) %>% 
  summarize(value = mean(value)) %>% 
  filter(day >= '2020-07-01' &
         day < '2020-08-01'
         ) %>% 
  collect() %>% 
  mutate(Lokalitet = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>% 
  mutate(data_type = factor(data_type, levels = )) 
##Why can't I set the levels in the pipe?
levels(tt$data_type) <- c("Duggpunkt", "Luftfuktighet", "Temperatur", "Lys")

tt %>%
  ggplot(.) +
  geom_line(aes(x = day, y = value, color = Lokalitet)) +
  facet_wrap(~data_type,
             scales = "free") +
  scale_color_nina(name = "Lokalitet") +
  xlab("Dag") +
  ylab("") +
  theme(legend.key.height = unit(0.5, "cm"))
  
```

There seems to be two clusters of relative humidity values. **In this updated data**, we see that only the Semi-nat_1 logger is off. That is the only one where we didn't download the data through Hobolink. So it seems that something happens when we run the data through hobolink, compared to when we download the data in a csv file directly from the logger. Must talk to Hobo about this.


```{r}
location_groups <- tibble(location = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)),
                          location_group = c("Oslo", "Lillehammer", "Lillehammer", "Lillehammer",
                                             "Oslo", "Lillehammer", "Lillehammer", "Lillehammer",
                                             "Oslo", "Oslo", "Oslo", "Oslo", "Oslo", "Lillehammer",
                                             "Lillehammer", "Lillehammer", "Oslo", "Lillehammer",
                                             "Oslo", "Lillehammer"))
```


```{r dew_point_group}
logger_data %>% 
    filter(day >= '2020-10-01' &
         day < '2020-11-01') %>% 
  group_by(location, data_type) %>% 
  summarize(value = mean(value)) %>% 
  filter(data_type == "dew") %>% 
  collect() %>% 
  mutate(location = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>% 
    left_join(location_groups,
            by = c("location" = "location")) %>% 
  ggplot(.) +
  geom_bar(aes(x = location, y = value, fill = location_group), stat = "identity") +
  scale_fill_nina(name = "Lokalitets-gruppe") +
  xlab("Lokalitet") +
  ylab("Middels duggpunkt i Juli") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```

Plot all but the Semi-nat_1 data.


```{r avg_logger_july_no_semi-nat_1}
tt <- logger_data %>% 
  filter(location != "Semi-nat_1") %>% 
  group_by(location, day, data_type) %>% 
  summarize(value = mean(value)) %>% 
  filter(day >= '2020-07-01' &
         day < '2020-08-01'
         ) %>% 
  collect() %>% 
  mutate(Lokalitet = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>% 
  mutate(data_type = factor(data_type, levels = )) 
##Why can't I set the levels in the pipe?
levels(tt$data_type) <- c("Duggpunkt", "Luftfuktighet", "Temperatur", "Lys")

tt %>%
  ggplot(.) +
  geom_line(aes(x = day, y = value, color = Lokalitet)) +
  facet_wrap(~data_type,
             scales = "free") +
  scale_color_nina(name = "Lokalitet") +
  xlab("Dag") +
  ylab("") +
  theme(legend.key.height = unit(0.5, "cm"))
  
```


<!-- What happens if we only take the day time values? -->
<!-- ```{r dew_point_group_day} -->
<!-- logger_data %>%  -->
<!--       filter(day >= '2020-07-01' & -->
<!--          day < '2020-08-01') %>% -->
<!--       collect() %>%  -->
<!--   mutate(time = as_hms(date)) %>%  -->
<!--   filter(time >= as_hms("12:00:00") & -->
<!--            time < as_hms("18:00:00")) %>%  -->
<!--   filter(data_type == "dew") %>%  -->
<!--   group_by(location) %>%  -->
<!--   summarize(value = mean(value)) %>%  -->
<!--   mutate(location = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>%  -->
<!--     left_join(location_groups, -->
<!--             by = c("location" = "location")) %>%  -->
<!--   ggplot(.) + -->
<!--   geom_bar(aes(x = location, y = value, fill = location_group), stat = "identity") + -->
<!--   scale_fill_nina(name = "Location group") + -->
<!--   xlab("Location") + -->
<!--   ylab("Average dew point in July") + -->
<!--   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  -->

<!-- ``` -->

<!-- Check correspondance with altitude -->
<!-- ------------ -->

<!-- ```{r} -->
<!-- location_alt_q <- " -->
<!-- SELECT location, avg(ST_Z(geom)) avg_altitude -->
<!-- FROM traps.trap_locations -->
<!-- WHERE elev_data IS TRUE -->
<!-- GROUP BY location -->
<!-- " -->
<!-- location_alt <- dbGetQuery(con, location_alt_q) -->

<!-- ``` -->

<!-- ```{r dew_point_alt} -->
<!-- logger_data %>%  -->
<!--     filter(day >= '2020-10-01' & -->
<!--          day < '2020-11-01') %>%  -->
<!--   group_by(location, data_type) %>%  -->
<!--   summarize(value = mean(value)) %>%  -->
<!--   filter(data_type == "dew") %>%  -->
<!--   collect() %>%  -->
<!--   mutate(location = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>%  -->
<!--     left_join(location_groups, -->
<!--             by = c("location" = "location")) %>%  -->
<!--   left_join(location_alt, -->
<!--             by = c("location" = "location")) %>%  -->
<!--   filter(!is.na(avg_altitude)) %>%  -->
<!--   ggplot(.) + -->
<!--   geom_bar(aes(x = avg_altitude, y = value, fill = location_group), stat = "identity") + -->
<!--   scale_fill_nina(name = "Lokalitets-gruppe") + -->
<!--   xlab("Lokalitets-altitud") + -->
<!--   ylab("Middels duggpunkt i Juli") + -->
<!--   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  -->


<!-- ``` -->

<!-- The dew points seem to correspond quite well with the altitude,  -->

<!-- ```{r rh_point_alt} -->
<!-- logger_data %>%  -->
<!--     filter(day >= '2020-10-01' & -->
<!--          day < '2020-11-01') %>%  -->
<!--   group_by(location, data_type) %>%  -->
<!--   summarize(value = mean(value)) %>%  -->
<!--   filter(data_type == "rh") %>%  -->
<!--   collect() %>%  -->
<!--   mutate(location = factor(location, levels = c(paste0("Semi-nat_", 1:10), paste0("Skog_", 1:10)))) %>%  -->
<!--     left_join(location_groups, -->
<!--             by = c("location" = "location")) %>%  -->
<!--   left_join(location_alt, -->
<!--             by = c("location" = "location")) %>%  -->
<!--   filter(!is.na(avg_altitude)) %>%  -->
<!--   ggplot(.) + -->
<!--   geom_bar(aes(x = avg_altitude, y = value, fill = location_group), stat = "identity") + -->
<!--   scale_fill_nina(name = "Lokalitets-gruppe") + -->
<!--   xlab("Lokalitets-altitud") + -->
<!--   ylab("Middels duggpunkt i Juli") + -->
<!--   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  -->


<!-- ``` -->



Sampling time analysis
========
Time to look at the impact of sampling time, 2 vs 4 weeks.


First off, simple plot of species richness against days sampled.

```{r spec_rich_vs_days_coll}
dat %>% 
  group_by(locality, 
           trap,
           trap_type, 
           empty_week,
           weeks_sampled,
           days_collected) %>% 
  summarise(no_spec = n_distinct(order, family, genus, species)) %>% 
  ggplot(aes(x = days_collected, y = no_spec)) +
  geom_point(aes(color = trap_type)) +
  scale_color_nina(name = "Felletype") +
  ylab("Antall taksa") +
  xlab("Innsamlingsdager")

```

Not easy to see any patterns here, since there is such variation in the catches throughout the season. We need to sum up the catches in the 2 2-week periods and compare to the single 4 week periods.

We first identify which of the 2-week samples that should be aggregated.
```{r}
#4-week "target" to agg to
week_4 <- dat %>% 
  filter(weeks_sampled == 4) %>% 
  select(locality,
         empty_week) %>% 
  distinct() %>% 
  mutate(agg_id = 1:n())

#add 2 weeks earlier
earlier_2_weeks <- week_4 %>% 
  select(locality,
         agg_id) %>% 
  mutate(empty_week = week_4$empty_week - 2)

which_to_agg <- week_4 %>% 
  bind_rows(earlier_2_weeks)

#add aggregation ids to 2-week samplings
to_agg <- dat %>% 
  filter(weeks_sampled == 2) %>% 
  left_join(which_to_agg,
            by = c("locality" = "locality",
                   "empty_week" = "empty_week"))
```

Check which didn't get an aggregation
```{r}
to_agg %>% 
  filter(is.na(agg_id)) %>% 
  select(locality,
         empty_week,
         trap) %>% 
  distinct() %>% 
  arrange(empty_week,
          locality) %>% 
  print(n = Inf)
```

The samples from week 24 are expected, since not all traps where set up early enough to get 2 emptyings by then.

```{r}
dat %>% 
  filter(locality == 'Semi-nat_2' &
           empty_week == 34) %>% 
  select(locality,
         empty_week,
         date_emptied,
         weeks_sampled,
         trap,
         GenlabID) %>% 
  distinct()
```

But Semi-nat_1, week_32, trap MF2 should be there. Semi-nat_5 as well. Seems that these haven't been run yet. I'll remove these for the comparison.



```{r}
#Sum the genetic reads to combine the two weeks
agg_data_2_week <- to_agg %>% 
  filter(!is.na(agg_id)) %>% 
  group_by(locality,
           trap_type,
           trap,
           agg_id,
           species_latin) %>% 
  summarise(value = sum(value)) %>% 
  ungroup() 

#Add the week information for the second sampling
agg_data_2_week <- agg_data_2_week %>% 
  left_join(week_4,
            by = c("agg_id" = "agg_id",
                   "locality" = "locality")) %>% 
  mutate(weeks_sampled = 2) %>% 
  select(locality,
         empty_week,
         weeks_sampled,
         trap_type,
         trap,
         species_latin,
         value)


#Select only the original 4 week samplings
data_4_week <- dat %>% 
  filter(weeks_sampled == 4) %>% 
    select(locality,
           empty_week,
           weeks_sampled,
           trap_type,
           trap,
           species_latin,
           value) 

#Add the original 4 weeks to the summarized 2 2-week samplngs
week_comp_data <- agg_data_2_week  %>% 
  bind_rows(data_4_week)

```

Summarise the number of species in both sample lengths.

```{r}
no_species_weeks <- week_comp_data %>% 
  group_by(locality,
           empty_week,
           weeks_sampled,
           trap_type) %>% 
  summarise(no_species = n())
```


```{r}
#Check comparison
no_species_weeks %>% 
  group_by(locality,
           empty_week,
           trap_type) %>% 
  summarise(no_comp = n()) %>% 
  print(n = Inf)
```

Make a comparison table.
```{r}
comp_species_week <- no_species_weeks %>% 
  pivot_wider(id_cols = c(locality, 
                          empty_week, 
                          trap_type),
              names_from = weeks_sampled,
              values_from = no_species,
              names_prefix = "sample_time_") %>% 
  mutate(diff_sample_time = sample_time_4 - sample_time_2,
         prc_diff_sample_time = (sample_time_4 - sample_time_2) / sample_time_2)
```

Quality check of a few rows. Seems OK.

```{r}
dat %>% 
  filter(locality == "Semi-nat_1" &
           (empty_week == 26 | empty_week == 28) &
           trap == "MF1"
         ) %>% 
  summarize(no_spec = n_distinct(species_latin))


dat %>% 
  filter(locality == "Semi-nat_5" &
           (empty_week == 22 | empty_week == 24) &
           trap == "MF1"
         ) %>% 
  summarize(no_spec = n_distinct(species_latin))

dat %>% 
  filter(locality == "Semi-nat_5" &
           (empty_week == 24 ) &
           trap == "MF2"
         ) %>% 
  summarize(no_spec = n_distinct(species_latin))

dat %>% 
  filter(locality == "Semi-nat_6" &
           (empty_week == 26 ) &
           trap == "MF2"
         ) %>% 
  summarize(no_spec = n_distinct(species_latin))

week_comp_data %>% 
  filter(locality == "Semi-nat_5" &
           (empty_week == 24 ) &
           trap == "MF2"
         ) %>% 
  summarize(no_spec = n_distinct(species_latin))

no_species_weeks %>% 
  filter(locality == "Semi-nat_5" & 
           empty_week == 24)

```



```{r richness_week_malaise, fig.cap = "Difference (g.) in cumulative species richness over 2 2-week samplings, compared to 1 4 week sampling. Catch in 1 Malaise trap per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
comp_species_week %>% 
  group_by(locality) %>% 
  filter(trap_type == "Malaise") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = diff_sample_time, color = locality)) +
  geom_line(aes(x = empty_week, y = diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("Diff. kumulativ 2 uker - 4 uker (no. taksa)") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) +
  theme(legend.key.height = unit(0.5, "cm"))


```

```{r prc_richness_week_malaise, fig.cap = "Percentage difference in cumulative species richness over 2 2-week samplings, compared to 1 4 week sampling. Catch in 1 Malaise trap per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
comp_species_week %>% 
  group_by(locality) %>% 
  filter(trap_type == "Malaise") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) + 
  geom_line(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("% Diff. kumulativ 2 uker - 4 uker") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) +
  theme(legend.key.height = unit(0.5, "cm"))


```

```{r richness_week_window, fig.cap = "Difference in cumulative species richness over 2 2-week samplings, compared to 1 4 week sampling. Catch in 1 Malaise trap per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
comp_species_week %>% 
  group_by(locality) %>% 
  filter(trap_type == "Vindu") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = diff_sample_time, color = locality)) +
  geom_line(aes(x = empty_week, y = diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("Diff. kumulativ 2 uker - 4 uker") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) 

```

```{r prc_richness_week_window, fig.cap = "Percentage difference in cumulative species richness over 2 2-week samplings, compared to 1 4 week sampling. Catch in 2 window traps per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
comp_species_week %>% 
  group_by(locality) %>% 
  filter(trap_type == "Vindu") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  geom_line(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("% Diff. kumulativ 2 uker - 4 uker") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) 

```

A similar comparison, but for biomass.
-------------

```{r}
biomass_agg_data_2_week <- to_agg %>% 
  filter(!is.na(agg_id)) %>% 
  select(locality,
         empty_week,
           trap_type,
           trap,
           agg_id,
         biomass = Toerrvekt_.g.) %>%
  distinct() %>% 
  arrange(locality,
          empty_week,
          trap) %>% 
  group_by(locality,
           trap_type,
           agg_id) %>% 
  summarise(empty_week = max(empty_week),
            biomass = sum(biomass)
            ) %>% 
  ungroup()  %>% 
  select(-agg_id)



biomass_data_4_week <- dat %>% 
  filter(weeks_sampled == 4) %>% 
  select(locality,
         empty_week,
         trap_type,
         trap,
         biomass = Toerrvekt_.g.) %>%
  distinct() %>% 
  arrange(locality,
          empty_week,
          trap) %>% 
    group_by(locality,
             empty_week,
           trap_type) %>% 
    summarise(empty_week = max(empty_week),
            biomass = sum(biomass)
            ) %>% 
  ungroup()

#Make comparison table
biomass_comp <- biomass_agg_data_2_week %>% 
  left_join(biomass_data_4_week,
            by = c("locality" = "locality",
                   "empty_week" = "empty_week",
                   "trap_type" = "trap_type"),
            suffix = c("_2_weeks", "_4_weeks")) %>% 
  mutate(diff_sample_time = biomass_4_weeks - biomass_2_weeks,
         prc_diff_sample_time = (biomass_4_weeks - biomass_2_weeks) / biomass_2_weeks)
  
  
```


```{r prc_biomass_week_malaise, fig.cap = "Percentage difference in cumulative dry weight over 2 2-week samplings, compared to 1 4 week sampling. Catch in 1 Malaise trap per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
biomass_comp %>% 
  group_by(locality) %>% 
  filter(trap_type == "Malaise") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  geom_line(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("% Diff. kumulativ 2 uker - 4 uker") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) +
  theme(legend.key.height = unit(0.5, "cm"))
```

We can do the same comparison with wet weight, althought we might expect that this will add some weight of the ethanol and the bottles, which will lead to bias when we sum these together for the 2 weeks. Anyway, this is the way it looks for wet weight.



```{r}
biomass_agg_data_2_week <- to_agg %>% 
  filter(!is.na(agg_id)) %>% 
  select(locality,
         empty_week,
           trap_type,
           trap,
           agg_id,
         biomass = Vaatvekt_.g.) %>%
  distinct() %>% 
  arrange(locality,
          empty_week,
          trap) %>% 
  group_by(locality,
           trap_type,
           agg_id) %>% 
  summarise(empty_week = max(empty_week),
            biomass = sum(biomass)
            ) %>% 
  ungroup()  %>% 
  select(-agg_id)



biomass_data_4_week <- dat %>% 
  filter(weeks_sampled == 4) %>% 
  select(locality,
         empty_week,
         trap_type,
         trap,
         biomass = Vaatvekt_.g.) %>%
  distinct() %>% 
  arrange(locality,
          empty_week,
          trap) %>% 
    group_by(locality,
             empty_week,
           trap_type) %>% 
    summarise(empty_week = max(empty_week),
            biomass = sum(biomass)
            ) %>% 
  ungroup()

#Make comparison table
biomass_comp <- biomass_agg_data_2_week %>% 
  left_join(biomass_data_4_week,
            by = c("locality" = "locality",
                   "empty_week" = "empty_week",
                   "trap_type" = "trap_type"),
            suffix = c("_2_weeks", "_4_weeks")) %>% 
  mutate(diff_sample_time = biomass_2_weeks - biomass_4_weeks,
         prc_diff_sample_time = (biomass_2_weeks - biomass_4_weeks) / biomass_2_weeks)
  
  
```


```{r prc_wet_weight_week_malaise, fig.cap = "Percentage difference in cumulative wet weight over 2 2-week samplings, compared to 1 4 week sampling. Catch in 1 Malaise trap per period. Negative values indicate that 4-week samplings gave lower insect species richness than the two 2-week samplings."}
biomass_comp %>% 
  group_by(locality) %>% 
  filter(trap_type == "Malaise") %>% 
  ggplot(.) +
  geom_point(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  geom_line(aes(x = empty_week, y = prc_diff_sample_time, color = locality)) +
  scale_color_nina(name = "Lokalitet") +
  ylab("% Diff. kumulativ 2 uker - 4 uker") +
  xlab("Tømmingsuke, 4 ukerstømming") +
  geom_hline(yintercept = 0) +
  theme(legend.key.height = unit(0.5, "cm"))

```


Compare the ethanol to the propglykol window traps
===============

```{r}
liquid_comp <- dat %>% 
  filter(trap_type == "Vindu") %>% 
  group_by(locality,
           trap,
           liquid,
           weeks_sampled,
           empty_week,
           sample_id
           ) %>% 
  summarise(no_species = n_distinct(species_latin))

liquid_comp
```

```{r}

liquid_mod_0 <- glmer(no_species ~ liquid * weeks_sampled + (1|empty_week) + (1 | locality), 
                      family = "poisson",
                      data = liquid_comp)

isSingular(liquid_mod_0)

summary(liquid_mod_0)
```

I struggle to model this with all the random effect I would like. I'm not able to include a id level random effect to account for overdispersion for example. Anyway, the results give an indication of the differences in the data.

```{r}
liquid_pred <- ggpredict(liquid_mod_0,
                         terms = c("liquid", "weeks_sampled"))
```


```{r liquid_window_plot, fig.cap = "Modelled difference in number of taxa found in window traps, in 2/4 weeks sampling and ethanol/propylen-glycole liquid."}
plot(liquid_pred) +
  scale_color_nina(name = "Antall uker") +
  ggtitle("") +
  ylab("Antall taksa") +
  xlab("Væske i vindusfelle")
```


